---
title: "The REAL McCOIL Rcpp package"
author: "OJ Watson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        toc: yes
        css: knitr.css
        code_folding: show
        fig_caption: yes
        theme: readable
        keep_md: yes
        toc_float: yes

---
    
## Overview
    
1. Description of aims of package
2. Testing Rcpp package
3. Reducing SNP sites in Uganda study



## 1. Data from Citygrapher 1 

*McCOILR* is simply an Rcpp implementation of [THEREALMcCOIL] (https://github.com/Greenhouse-Lab/THEREALMcCOIL),
which was written solely to make running the software easier within the cluster
framework I use. All rights refer to the writers of the original c code.

The package can be installed as follows, assuming devtools has been installed. 

```{r Load package, include=TRUE, message = FALSE, warning = FALSE,cache=TRUE}

## first let's install the package
devtools::install_github("OJWatson/McCOILR")

## Load the package
library(McCOILR)

```

## 2. Testing Rcpp package

The package carries out the same 2 R functions as before, which are demonstrated below: 

```{r Test Package, include=TRUE, message = FALSE, warning = FALSE,cache=TRUE}

## categorical test

# read in demo data and view it
data0 = read.table(system.file("extdata","cat_input_test.txt",package="McCOILR"), head=T)
data=data0[,-1]
rownames(data)=data0[,1]

# view the heterozygosity calls
head(data)

# create results directory and run analysis
dir.create(path = "cat_output")
out_cat <- McCOIL_categorical(data,maxCOI=25, threshold_ind=20, threshold_site=20, totalrun=1000, burnin=100, M0=15, e1=0.05, 
                   e2=0.05, err_method=3, path="cat_output", output="output_test.txt" )


## proportional test

# read in demo data and view it
dataA1i = read.table(system.file("extdata","prop_dataA1_test.txt",package="McCOILR"), head=T)
dataA2i = read.table(system.file("extdata","prop_dataA2_test.txt",package="McCOILR"), head=T)
dataA1= dataA1i[,-1]
dataA2= dataA2i[,-1]
rownames(dataA1)= dataA1i[,1]
rownames(dataA2)= dataA2i[,1]

# view the read counts
head(dataA1)

# create results directory and run analysis
dir.create(path="prop_output")
out_prop <- McCOIL_proportional(dataA1, dataA2, maxCOI=25, totalrun=5000, burnin=100, M0=15, epsilon=0.02, err_method=3, 
                                path="prop_output", output="output_test.txt" )


```

The R functions now return the summary outputs, just for ease of looking at the 
results:

```{r View output, include=TRUE, message = FALSE, warning = FALSE,cache=TRUE}

## view summary data.frame for categorical
str(out_cat)

## Have a look at the categorical output COI distribution
hist(as.numeric(as.character(out_cat$mean[out_cat$CorP=="C"])),main = "Categorical Mean COI", xlab="COI")


## view summary data.frame for proportional
str(out_prop)

## Have a look at the proportional output COI distribution
hist(as.numeric(as.character(out_prop$mean[out_cat$CorP=="C"])),main = "Proportional Mean COI", xlab="COI")

```

## 3. Reducing SNP sites in Uganda study

The main reason the package was created was so that the software could be easily
run on my cluster framework in my department. This was because I wanted to see
what effect reducing the number of SNP locations would have in terms of 
increasing or decreasing the estimated COI. 

```{r Cluster Work, include=TRUE, message = FALSE, warning = FALSE,  eval=F, echo=T}

# Read in Uganda dataset
snps <- read.csv(system.file("extdata","SNP.txt",package="McCOILR"),sep="\t")

# let's just look at Kihihi
kihihi_data <-  snps[snps$location=="Kihihi",]

# create 100 sets of 24, 48 and 96 SNP site samples
barcode24 <- replicate(sample(3:105,size = 24,replace = FALSE),n = 100)
barcode48 <- replicate(sample(3:105,size = 48,replace = FALSE),n = 100)
barcode96 <- replicate(sample(3:105,size = 96,replace = FALSE),n = 100)

## The following is code to run the categorical analysis on our cluster but is 
## just looping through the above barcode sets and running the categorical analysis
## on each one

## --------------------------------------------------------------------------##

workdir <- "M:/OJ/McCOILR_Results"
didehpc::didehpc_config_global(workdir=workdir,
                               credentials="C:\\Users\\Oliver\\.smbcredentials",
                               temp=didehpc::path_mapping("tmp","T:","//fi--didef3.dide.ic.ac.uk/tmp","T:"),
                               home=didehpc::path_mapping("OJ","M:","//fi--didef3.dide.ic.ac.uk/Malaria","M:"))
didehpc::web_login()
root <- file.path(workdir, "contexts")
packages.vector <- c("Rcpp","McCOILR")
context::context_log_start()
ctx <- context::context_save(root,
                             packages = packages.vector,
                             package_sources= provisionr::package_sources(github=c("OJWatson/McCOILR")))

config <- didehpc::didehpc_config(use_workers = TRUE)
obj <- didehpc::queue_didehpc(ctx, config = config)

paramList <- list()
length(paramList) <- 300

for(i in 1:300){

    if(i <=100){
        data <- kihihi_data[,barcode24[,i]+2]
        output <- paste0("Kihihi_24BarcodeTest_50maxCOI/24_",i,".txt")
    } else if (i > 100 & i <= 200) {
        data <- kihihi_data[,barcode48[,i-100]+2]
        output <- paste0("Kihihi_48BarcodeTest_50maxCOI/48_",i-100,".txt")
    } else {
        data <- kihihi_data[,barcode96[,i-200]+2]
        output <- paste0("Kihihi_96BarcodeTest_50maxCOI/96_",i-200,".txt")
    }
    

paramList[[i]] <- (list(data = data, maxCOI=50, threshold_ind=20, 
                       threshold_site=20, totalrun=100000, burnin=100,
                       M0=15, e1=0.05, e2=0.05, err_method=3, 
                       path="M:/OJ/McCOILR_Results",
                        output=output))

}

dir.create(paste0(workdir,"/Kihihi_24BarcodeTest_50maxCOI"))
dir.create(paste0(workdir,"/Kihihi_48BarcodeTest_50maxCOI"))
dir.create(paste0(workdir,"/Kihihi_96BarcodeTest_50maxCOI"))

workers <- obj$submit_workers(50)
grp <- queuer::qlapply(X = paramList[1:100],obj = obj, timeout=0, 
                       FUN = function(x){return(McCOIL_categorical(data = x$data,maxCOI = x$maxCOI, threshold_ind = x$threshold_ind,
                                                                   threshold_site = x$threshold_site, totalrun = x$totalrun, 
                                                                   burnin = x$burnin, M0 = x$M0, e1 = x$e1, e2 = x$e2, 
                                                                   err_method = x$err_method, path = x$path, output = x$output)
                                                )},
                       name = "24_barcode")

## --------------------------------------------------------------------------##

```

The code above is executing the categorical method on 100 different randomly chosen
sets of SNP sites, with sets equal to 24, 48, 96, with 100,000 MCMC iterations 
and a maximum possible COI of 50.

The summary results of these MCMC iterations has been copied into the tutorials folder, 
so the following analysis can be verified. Below I'm simply showing the overall 
distribution for mean COI for the sets of 24, 48 and 96 SNP sites. 

```{r Plot pseudo-steiner-again, include=TRUE, message = FALSE, warning = FALSE, cache=TRUE}

# storage for the 24, 48 and 96 site datasets
twenty4 <- list()
fourty8 <- list()
ninety6 <- list()

for(i in 1:100){
    
  twenty4[[i]] <- read.csv(paste0("Kihihi_24_BarcodeTest_50maxCOI/24_",i,".txt_summary.txt"),sep="\t")
  fourty8[[i]] <- read.csv(paste0("Kihihi_48_BarcodeTest_50maxCOI/48_",i,".txt_summary.txt"),sep="\t")
  ninety6[[i]] <- read.csv(paste0("Kihihi_96_BarcodeTest_50maxCOI/96_",i,".txt_summary.txt"),sep="\t")
  
  

}

## Grab the estimated mean COI and lets plot these
mean24 <- lapply(twenty4,function(x){return(as.numeric(as.character(x$mean[x$CorP=="C"])))})
mean48 <- lapply(fourty8,function(x){return(as.numeric(as.character(x$mean[x$CorP=="C"])))})
mean96 <- lapply(ninety6,function(x){return(as.numeric(as.character(x$mean[x$CorP=="C"])))})

par(mfrow=c(1,3))
hist(as.numeric(unlist(mean24)),main="24 Sites",xlab="COI")
hist(as.numeric(unlist(mean48)),main="48 Sites",xlab="COI")
hist(as.numeric(unlist(mean96)),main="96 Sites",xlab="COI")


```

Initially it seemed to me that the fewer SNP sites was causing a larger COI to be
estimates. However, upon further inspection of the chains it could be seen that 
using fewer SNP sites resulted in MCMC chains failing to converge more often, which
gave spurious results.

```{r Plot example chains, include=TRUE, message = FALSE, warning = FALSE, cache=TRUE}

## First let's grab one MCMC output with 24 sites
MCMC_24 <- read.table("M:/OJ/McCOILR_Results/Kihihi_24BarcodeTest_50maxCOI/24_99.txt",head=F)

## First let's grab one MCMC output with 96 sites
MCMC_96 <- read.table("M:/OJ/McCOILR_Results/Kihihi_96BarcodeTest_50maxCOI/96_99.txt",head=F)

## We'll save these in the tutorials folder too for reproducibility
save(list=c("MCMC_24","MCMC_96"),file = "MCMC24&96.RData")

## Now let's plot some random COI estimate chains

# for MCMC24 we will plot 9 random COI and allele frequency chains with some thinning
par(mfrow=c(3,3))
for(i in floor(seq(2,length(MCMC_24)-2,length.out = 9))){
    plot(MCMC_24[[i]][seq(1,100000,50)],type="l",
         ylab = "COI", main = twenty4[[99]]$name[i-1])
}

# and the same for MCMC96
par(mfrow=c(3,3))
for(i in floor(seq(2,length(MCMC_96)-2,length.out = 9))){
    plot(MCMC_96[[i]][seq(1,100000,50)],type="l",
         ylab = ifelse(length(grep(pattern = "assay",as.character(ninety6[[99]]$name[i-1]))==1),"Freq","COI"),
         main = ninety6[[99]]$name[i-1])
}


```

We can also see that when we use the 96 SNP sites, the overall mean COI is equal to
2.60, with a median equal to 2 which agrees very well with the analysis presented of 
the categorical method for Kihihi within the supplementary material for THE REAL
McCOIL, whereas for 24 sites these do not agree:

```{r Mean COI, include=TRUE, message = FALSE, warning = FALSE, cache=TRUE}

## Mean for 24
print( paste0("Mean for 24 = ", mean(as.numeric(unlist(mean24)))))

## Median for 24
print( paste0("Median for 24 = ", median(as.numeric(unlist(mean24)))))

## Mean for 96
print( paste0("Mean for 96 = ", mean(as.numeric(unlist(mean96)))))

## Median for 96
print( paste0("Median for 24 = ", median(as.numeric(unlist(mean96)))))
```

As such it makes sense in the future when using THE REAL McCOIL on new datasets
to use more sites within computational reason that display both polymorphic and 
intermediate/high allele frequencies in multiple populations, such as the 128 SNPs 
used within the publication attached to THE REAL McCOIL. 

In addition I will be adding functionality to McCOILR to pipeline chain convergence
assessment, such that when x number of repetitions on the same dataset do not yield 
convergent results using Gelman and Rubin convergence diagnostics additional repetitions
are automated until a convergent result is achieved. 